\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{geometry}
\geometry{margin=1in}

\title{\textbf{CHITTA: The Cognitive Horizon (Fibonacci Neural Substrate)}}
\author{Neoned Developer \\ Independent Researcher \\ \texttt{[REDACTED]}}
\date{February 2026}

\begin{document}

\maketitle

\begin{abstract}
We introduce \textbf{CHITTA} (Cognitive Horizon of Infinite Time \& Telemetric Awareness), formally known as the \textbf{Fibonacci Neural Substrate (FNS)}. This architecture enables true self-awareness in autonomous AI agents through hierarchical micro-processing units operating on fractal-compressed temporal data—mirroring the ancient concept of \textit{Chitta} (mind-stuff) accumulating \textit{Samskaras} (impressions) over \textit{Kalpa} (cycles of time). Unlike existing agent monitoring systems that treat observation and action as separate concerns, CHITTA unifies sensory input (\textit{Prana}), memory compression (\textit{Smriti}), constitutional reasoning (\textit{Dharma}), and interrupt-driven responses into a continuous neural substrate.
\end{abstract}

\section{Introduction}
Modern autonomous AI agents face a critical paradox: as they grow more capable, they become less observable to themselves. Current monitoring approaches treat agents as black boxes instrumented with external metrics, rather than as integrated systems capable of perceiving their own state. This externalized observation model fails when:
\begin{enumerate}
    \item \textbf{Freeze Detection:} A 5.5-hour agent freeze goes undetected.
    \item \textbf{Resource Exhaustion:} Out-of-memory crashes occur without preemptive action.
    \item \textbf{Constitutional Drift:} Agents violate core directives without self-correction.
\end{enumerate}

We propose that \textbf{self-awareness is not an emergent property but an architectural requirement}—a substrate that must be explicitly designed.

\section{Fibonacci Neural Substrate Architecture}
The Fibonacci Neural Substrate consists of five interconnected layers, mirroring biological nervous systems (L1-Brainstem to L5-Superego).

\subsection{Fibonacci Fractal Pyramid (Layer 3)}
Traditional fixed-resolution sampling (e.g., Prometheus) discards temporal structure. We instead use Fibonacci-based hierarchical compression. Given atomic samples $S$ at 5-second intervals, we partition $S$ into windows of size $F(3), F(4), F(5), ..., F(k)$. For each window, we compute OHLC (Open, High, Low, Close) plus mean $\mu$.

\subsubsection{Efficiency}
For 31 metrics sampled at 5s over 30 days, FFP achieves a 13,824:1 compression ratio, reducing 15.8 MB of raw data to 1.6 KB while maintaining 8ms query latency.

\section{Implementation: NEO Case Study}
\textbf{NEO} (Neural Evolutionary Organism) is a 4-tier autonomous agent running on WSL2 with SmolLM3-3B local inference. Phase 0 implementation of Heartbeat and RAM Guardian prevented 3 freeze incidents in the first 24 hours.

\section{Conclusion}
Self-awareness is not emergent magic—it's engineered architecture. By unifying fractal compression and hierarchical micro-processing, FNS enables agents to perceive, remember, and ethically regulate their own behavior in real-time.

\begin{thebibliography}{9}
\bibitem{jacquin1992} Jacquin, A. E. (1992). Image coding based on a fractal theory of iterated contractive image transformations.
\bibitem{anthropic2022} Anthropic (2022). Constitutional AI: Harmlessness from AI Feedback.
\bibitem{sutton1999} Sutton, R. S., et al. (1999). Between MDPs and semi-MDPs: A framework for temporal abstraction.
\end{thebibliography}

\end{document}
